### 2018-11-13 Sabanci_talk/Sabanci_talk.tex
FINAL version

 \begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  \textcolor{white}{{\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$}
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, GRBs, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E}=L_\text{GW} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  {\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\animategraphics[loop,controls,height=2.9cm]{12}{Circ_Orb-}{0}{20}};
    %\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-1.3cm)}]current page.center) {$\mu=m_1 m_2/M^2,\ x(t) = r \cos\Omega t, \ y(t) = r \sin\Omega t $};
    \onslide<3->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(2.25cm,-2.2cm)}]current page.center) {$ Q^{ij} =  \mu\, x^i x^j, \qquad \boxed{\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}r^4 \Omega^6}$};
    \onslide<4->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2.5cm,-3.2cm)}]current page.center) {Enhancement due to eccentricity \bl{$e$}};
    \node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-4cm)}]current page.center) {$ \bl{\f{\left(1 + \f{73}{24}e^2+ \f{37}{96} e^4 \right)}{(1-e^2)^{7/2}}}$};
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}{Gravitational waves from BNS}
  TT gauge, multipole, slow-velocity expansions give 
  \begin{small}
    \[
    h_{ij}^\text{TT} = \tfrac{2}{D}\tfrac{G}{c^4} \ddot{Q}_{ij}^\text{TT}\hspace{6cm}
    \]
    %
    \vspace{-5mm}
    \uncover<2->
    {
      \begin{align*}
      h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi(t)],\hspace{5cm}\\
      h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi(t)],
      \end{align*}
    }
  \end{small}
  %
  %
  \uncover<2->{$\iota =\,$ orbital \bl{inclination}}
  \vspace{-3mm}
  %
  \uncover<3->
  {
    \begin{small}
      \begin{align*}
      h_c(t) %&=\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3}\nn\\
      &\approx 7.5\times \bl{10^{-24}} \left(\f{100\,\text{Mpc}}{D}\right)\left(\f{M_c}{1.219 M_\odot}\right)^{5/3}\left(\frac{\red{f}}{10\,\text{Hz}}\right)^{\red{2/3}} \\
      \Phi(t) &=\int_{t_i}^t dt' \omega(t') + \text{PN}
      \end{align*}
    \end{small}
    \\
  }
  %
  \uncover<4->
  {
    Frequency increases as $f^{11/3}$\\
    Amplitude increases as $f^{10/3} $
  }  
  %
  %
  \uncover<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.5cm,1.2cm)}]current page.center) {\includegraphics[height=3cm]{figs/orbit_inc.png}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(3.7cm,-0.1cm)}]current page.center) {\textcolor{white}{\url{www.handprint.com/ASTRO/bineye3.html}}};
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/white_box.png}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\textcolor{white}{Chirp!}}};
    \end{tikzpicture}
  }
  \uncover<5->
  {
      \begin{tikzpicture}[overlay,remember picture]
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/Chirp.pdf}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\red{Chirp!}}};
    \end{tikzpicture}
  }  
\end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    \onslide<2->
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      \onslide<3->
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(L,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\includegraphics[height=2.39cm]{figs/white_box.png}};
      \end{tikzpicture}
          \begin{small}
      \begin{align*}
	\w{s_{12} }& \w{=\int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} }\\
	   &\w{ \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] }\\
      \w{s_{13}} &\w{\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]}
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    %
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      %
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(L,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\animategraphics[loop,controls,height=2.39cm]{12}{LIGO_int-}{0}{102}};
      \end{tikzpicture}
    \onslide<2->      
      \begin{small}
      \begin{align*}
	s_{12} &= \int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} \\
				    & \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] \\
      s_{13} &\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

@@

\begin{frame}{Detection criteria}{Optimal Signal-to-Noise Ratio ({\bf SNR})}
  \vspace{-5mm}
  \begin{footnotesize} 
  \[
  \rho=\left[ \int_0^\infty d\ln f\, \f{|\red{2\tilde{h}(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  \]
  \\
  \end{footnotesize}
  $\sqrt{S_n(f)}\equiv$ {\it amplitude spectral density} ({\bf ASD}) or  detector noise
  \\
  \onslide<2->
  {
    \vspace{3mm}
    Recall, $\tilde{h}(f) \propto f^{-7/6}$ so $\tilde{h}(f) \sqrt{f} \propto \red{f^{-2/3}}$
    \begin{footnotesize} 
    \[
    \rho^2 =  (2A)^2 h_0^2 |Q(\theta,\phi,\psi,\iota)|^2 \int_0^{\bl{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)} 
    \]
    \end{footnotesize}
  }
  %
  \onslide<3->
  RMS \red{average} over the angles 
  \begin{small}
  $ \langle \ldots \rangle \equiv \int \sin\theta d\theta d\phi \int d\psi \int\sin\iota d\iota $
  \end{small}
  \begin{footnotesize} 
  \[
  \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\llcorner = \f{4}{25}, \qquad \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\triangle = \f{9}{25}
  \]
  \end{footnotesize}
  %
  Sources will have random $\{\theta,\phi,\psi,\iota\}$ \\
  \begin{footnotesize} 
  \[
  0 \le Q(\theta,\phi,\psi,\iota) \le \underbrace{Q_\text{max}(\theta,\phi,\psi,\iota)}_{\theta=\tfrac{\pi}{2},\psi=0,\iota=0}
  \]
  \end{footnotesize}
  % 
\end{frame}

\begin{frame}{L-IFO vs. $\bigtriangleup$-IFO Networks}
\red{Three L}-IFOs, e.g., LIGO L1, H1 + Virgo
%
  \begin{footnotesize} 
  \begin{align*}
  \rho_{\text{tot},\red{\llcorner}} &= \left[{\sum_{i=1}^3 \rho^2_i}\right]^{1/2} =\left[ \rho_h^2 + \rho_m^2 + \rho^2_l \right]^{1/2} 
	  = {\rho_c\left[\tilde{A}_h^2 + \tilde{A}_m^2 + \tilde{A}_l^2 \right]^{1/2}}\\
  \rho_c &\equiv 2 h_0 \left[ \int_0^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2}, \quad \left\{ \tilde{A}_h = \tfrac{4}{5}A,  \tilde{A}_m =\tfrac{2}{5}A,  \tilde{A}_l  < \tfrac{\tilde{A}_m}{2}\right\}
  \end{align*}
  \end{footnotesize}
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.5cm,minimum height=0.8cm] (b) at (8.8,1.9){};
  \end{tikzpicture}
  \\
  %\vspace{2mm}
  \onslide<2->
  {
    $\bigtriangleup$ = \red{Three V-}IFOs: Einstein Telescope
    \begin{footnotesize} 
    \[
    \rho_{\text{tot},\bigtriangleup} = \left[\sum_{A=1}^3 \left(\rho^A\right)^2\right]^{1/2}  =  A \rho_c \,|Q(\theta,\phi,\psi,\iota)|_\bigtriangleup %\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \,
    \]
    \end{footnotesize}
    \\
  }
  \onslide<3->
  Use \bl{RMS-averaged} $|Q|^2$
  \begin{footnotesize}
  \begin{align*}
  \rho_{\text{tot},\bigtriangleup} %&= \left[ \int_0^\infty d\ln f\, \f{|{2\tilde{H}_\bigtriangleup(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  &=\f{6}{5}A\, h_0 \rho_c \\%\left[\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)}\right]^{1/2} \\
  \end{align*}
  \end{footnotesize}
  \\
  %where \begin{small}$\tilde{H}_\bigtriangleup(f)\equiv A\,  \langle |Q|^2_\bigtriangleup\rangle\, h_0 f^{-7/6}= \f{3}{5} A h_0 f^{-7/6}$ \end{small}\\
  \vspace{-5mm}
  Compare with three-L network's \begin{small}$\rho_\text{tot} \approx \tfrac{4}{5}A\, h_0 \left[ \int df \ldots \right]^{1/2} $ \end{small}
  \vspace{10mm}
  %
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.cm,minimum height=0.9cm] (b) at (-3.2,0.9){};
  \end{tikzpicture}
  %
  % 
\end{frame}

@@

  \begin{columns}
  \begin{column}{0.5\textwidth}  %%<--- here
      \begin{center}
      \onslide<3->\includegraphics[height=4cm]{figs/GW_GRB170817.jpg}
      \end{center}
  \end{column}
  \begin{column}{0.5\textwidth}
  %
  \onslide<6->
  {
    Assuming low spin priors
    \begin{align*}
     M&=2.73^{+0.04}_{-0.01} M_\odot \\
     M_c &=1.186^{+0.001}_{-0.001}M_\odot  \\ 
     D&=40.7^{+2.36}_{-2.36}\text{Mpc}
    \end{align*}
  }
  \onslide<7->
  {
    \red{Note:} $M_c$ has the best precision: 
   \begin{small}$${\dot{f}} = \f{96}{5}\pi^{8/3} \left(\tfrac{G M_c}{c^3}\right)^{5/3}\, {f^{11/3}}$$ \end{small}
  }

>>>>

 \begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  \textcolor{white}{{\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$}
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
  \end{tikzpicture}
\end{frame}

\begin{frame}{Making gravitational waves}{non-spherical, accelerated motion}
  \alert{Inspirals}, bumps on neutron stars, supernovae, the Big Bang. \\%  (BICEP 2).\\
  Gravitational-wave emission: \alert{quadrupole} radiation (leading-order)
  \begin{center} $ \boxed{\dot{E} = \f{G}{c^5} \langle (\partial_t^3 \bar{Q}_{ij})^2 \rangle}$\\
  \vspace{2mm}
  \begin{small}$ \quad \bar{Q}_{ij} = Q_{ij}-\frac{1}{3}\delta_{ij}Q^k_k, \quad Q^{ij} = \int \rho(t,{\mathbf x})x^i x^j d^3 x $ \end{small} \end{center}
  {\bf Binary systems:} 2 points masses in a circular orbit $M=m_1+m_2$
  \begin{tikzpicture}[overlay,remember picture]
    \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\animategraphics[loop,controls,height=2.9cm]{12}{Circ_Orb-}{0}{20}};
    %\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(-3.8cm,-2.7cm)}]current page.center) {\includegraphics[height=2.9cm]{figs/white_box.png}};
    \onslide<2->\node (t1 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-1.3cm)}]current page.center) {$\mu=m_1 m_2/M^2,\ x(t) = r \cos\Omega t, \ y(t) = r \sin\Omega t $};
    \onslide<3->\node (t2 )[anchor=center,scale=1,opacity=1] at ([shift={(2.25cm,-2.2cm)}]current page.center) {$ Q^{ij} =  \mu\, x^i x^j, \qquad \boxed{\dot{E} = \f{32}{5}\f{G\mu^2}{c^5}r^4 \Omega^6}$};
    %\onslide<4->\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2.5cm,-3.2cm)}]current page.center) {Enhancement due to eccentricity \bl{$e$}};
    %\node (t3 )[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-4cm)}]current page.center) {$ \bl{\f{\left(1 + \f{73}{24}e^2+ \f{37}{96} e^4 \right)}{(1-e^2)^{7/2}}}$};
  \end{tikzpicture}
\end{frame}

@@

\begin{frame}{Gravitational waves from BNS}
  %TT gauge, multipole, slow-velocity expansions give 
  %\begin{small}
   % \[
    %h_{ij}^\text{TT} = \tfrac{2}{D}\tfrac{G}{c^4} \ddot{Q}_{ij}^\text{TT}\hspace{6cm}
    %\]
    %
    %\vspace{-5mm}
    Recall $h_{\mu \nu} \approx A_{\mu\nu}\cos\left[\omega (t-z/c)\right]$ \\
    %  
    \begin{small}
    \uncover<2->
    {
      \begin{align*}
      A_{xx}&=h_+, A_{xy} = h_\times, \\
      h_+(t) &= h_c(t)\, \left(\f{1+\cos^2\iota}{2}\right)\, \cos[\Phi(t)],\hspace{5cm}\\
      h_\times(t) & =h_c(t)\,\cos\iota \sin[\Phi(t)],
      \end{align*}
    }
  \end{small}
  %
  %
  \uncover<2->{$\iota =\,$ orbital \bl{inclination}}
  \vspace{-3mm}
  %
  \uncover<3->
  {
    \begin{small}
      \begin{align*}
      h_c(t) %&=\f{4}{D}\left(\f{G M_c}{c^2}\right)^{5/3}\, \left(\f{\pi f(t)}{c}\right)^{2/3}\nn\\
      &\approx 7.5\times \bl{10^{-24}} \left(\f{100\,\text{Mpc}}{D}\right)\left(\f{M_c}{1.219 M_\odot}\right)^{5/3}\left(\frac{\red{f}}{10\,\text{Hz}}\right)^{\red{2/3}} \\
      \Phi(t) &=\int_{t_i}^t dt' \omega(t') + \text{PN}
      \end{align*}
    \end{small}
    \\
  }
  %
  \uncover<4->
  {
    Frequency increases as $f^{11/3}$\\
    Amplitude increases as $f^{10/3} $
  }  
  %
  %
  \uncover<2->
  {
    \begin{tikzpicture}[overlay,remember picture]
      \node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.5cm,1.2cm)}]current page.center) {\includegraphics[height=3cm]{figs/orbit_inc.png}};
      \node[anchor=center,scale=0.5,opacity=1] at ([shift={(3.7cm,-0.1cm)}]current page.center) {\textcolor{white}{\url{www.handprint.com/ASTRO/bineye3.html}}};
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/white_box.png}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\textcolor{white}{Chirp!}}};
    \end{tikzpicture}
  }
  \uncover<5->
  {
      \begin{tikzpicture}[overlay,remember picture]
      \node (img1 )[anchor=center,scale=1,opacity=1] at ([shift={(2.7cm,-3.2cm)}]current page.center) {\includegraphics[height=2.5cm]{figs/Chirp.pdf}};
      \node[anchor=center,scale=1.4,opacity=1] at ([shift={(1cm,-2.4cm)}]current page.center) {\href{https://www.youtube.com/watch?v=CH9JSqUfnSw}{\red{Chirp!}}};
    \end{tikzpicture}
  }  
\end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    \onslide<2->
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      \onslide<3->
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(0,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\includegraphics[height=2.39cm]{figs/white_box.png}};
      \end{tikzpicture}
          \begin{small}
      \begin{align*}
	\w{s_{12} }& \w{=\int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} }\\
	   &\w{ \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] }\\
      \w{s_{13}} &\w{\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]}
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

 \begin{frame}{GW Detectors: \red{Interferometers}}{Geodesic deviation}
    Two geodesics separated by $\xi^\mu$:
    %\[
    % \frac{D^2 \xi^\alpha}{D\tau^2} = -R^\alpha_{\beta \gamma\delta}\, \xi^\gamma u^\beta u^\delta 
    %\]
    $ \tfrac{d^2 \xi^i}{d\tau^2} = -R^i_{0j0} \xi^j\left(\tfrac{dx^0}{d\tau}\right)^2$ \hfill (detector frame)\\
    %\hspace{4cm} $\Longrightarrow \ddot{\xi^i} = -c^2 R^i_{0j0} \xi^j $ \hfill (detector time)\ \\
    \vspace{-3mm}
    \[
    \hspace{4.5cm} \boxed{\ddot{\xi^i} = \frac{1}{2} \ddot{h}_{ij}^\text{TT} \xi^j} \hspace{2.5cm} \text{(detector time)}
    \]
    %
    %
    %
    GW along $\mathbf{z}$: 
      \[
      h_{ij}^\text{TT} = \left(\begin{array}{cc}
				h_+ & h_\times \\
				h_\times & - h_+
				\end{array}\right) \cos\left(\omega\left( t-\tfrac{z}{c}\right)\right)
      \]
      %
      Consider mirrors at $(0,0)_1, (L,0)_2$ and $(0,L)_3$
      \begin{tikzpicture}[overlay,remember picture]
	%\node[anchor=center,scale=1,opacity=1] at ([shift={(2cm,-3.4cm)}]current page.center){\todo{Make sure the vid here works}} ;
	\node (img0 )[anchor=center,scale=1,opacity=1] at ([shift={(3.8cm,-2.8cm)}]current page.center) {\animategraphics[loop,controls,height=2.39cm]{12}{LIGO_int-}{0}{102}};
      \end{tikzpicture}
    \onslide<2->      
      \begin{small}
      \begin{align*}
	s_{12} &= \int_1^2 \tfrac{ds}{dx}dx = (x_2-x_1) \left[ 1 + h_+ \cos(\omega t)\right]^{1/2}\hspace{4cm} \\
				    & \approx L \left[ 1+\tfrac{1}{2} h_+ \cos\omega t \right] \\
      s_{13} &\approx L \left[ 1-\tfrac{1}{2} h_+ \cos\omega t \right]
      \end{align*}
      \end{small}
    %
    %
 \end{frame}

@@

\begin{frame}{Detection criterion}{Optimal Signal-to-Noise Ratio ({\bf SNR})}
  \vspace{-5mm}
  \begin{footnotesize} 
  \[
  \rho=\left[ \int_0^\infty d\ln f\, \f{|\red{2\tilde{h}(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  \]
  \\
  \end{footnotesize}
  $\sqrt{S_n(f)}\equiv$ {\it amplitude spectral density} ({\bf ASD}) or  detector noise
  \\
  \onslide<2->
  {
    \vspace{3mm}
    Recall, $\tilde{h}(f) \propto f^{-7/6}$ so $\tilde{h}(f) \sqrt{f} \propto \red{f^{-2/3}}$
    \begin{footnotesize} 
    \[
    \rho^2 =  (2A)^2 h_0^2 |Q(\theta,\phi,\psi,\iota)|^2 \int_0^{\bl{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)} 
    \]
    \end{footnotesize}
  }
  %
  \onslide<3->
  RMS \red{average} over the angles 
  \begin{small}
  $ \langle \ldots \rangle \equiv \int \sin\theta d\theta d\phi \int d\psi \int\sin\iota d\iota $
  \end{small}
  \begin{footnotesize} 
  \[
  \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\llcorner = \f{4}{25}, \qquad \langle |Q(\theta,\phi,\psi,\iota)|^2 \rangle_\triangle = \f{9}{25}
  \]
  \end{footnotesize}
  %
  %Sources will have random $\{\theta,\phi,\psi,\iota\}$ \\
  %\begin{footnotesize} 
  %\[
  %0 \le Q(\theta,\phi,\psi,\iota) \le \underbrace{Q_\text{max}(\theta,\phi,\psi,\iota)}_{\theta=\tfrac{\pi}{2},\psi=0,\iota=0}
  %\]
  %\end{footnotesize}
  % 
\end{frame}

\begin{frame}{L-IFO vs. $\bigtriangleup$-IFO Networks}
\red{Three L}-IFOs, e.g., LIGO L1, H1 + Virgo; or ($h$)igh, ($m$)edium, ($l$)ow
%
  \begin{footnotesize} 
  \begin{align*}
  \rho_{\text{tot},\red{\llcorner}} &= \left[{\sum_{i=1}^3 \rho^2_i}\right]^{1/2} =\left[ \rho_h^2 + \rho_m^2 + \rho^2_l \right]^{1/2} 
	  = {\rho_c\left[\tilde{A}_h^2 + \tilde{A}_m^2 + \tilde{A}_l^2 \right]^{1/2}}\\
  \rho_c &\equiv 2 h_0 \left[ \int_0^{{f_\text{ISCO}}} df \f{f^{-7/3}}{S_n(f)}  \right]^{1/2}, \quad \left\{ \tilde{A}_h = \tfrac{4}{5}A,  \tilde{A}_m =\tfrac{2}{5}A,  \tilde{A}_l  < \tfrac{\tilde{A}_m}{2}\right\}
  \end{align*}
  \end{footnotesize}
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.5cm,minimum height=0.8cm] (b) at (8.8,1.9){};
  \end{tikzpicture}
  \\
  %\vspace{2mm}
  \onslide<2->
  {
    $\bigtriangleup$ = \red{Three V-}IFOs: Einstein Telescope
    \begin{footnotesize} 
    \[
    \rho_{\text{tot},\bigtriangleup} = \left[\sum_{A=1}^3 \left(\rho^A\right)^2\right]^{1/2}  =  A \rho_c \,\langle|Q(\theta,\phi,\psi,\iota)|^2\rangle^{1/2}_\bigtriangleup %\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)} \,
    \]
    \end{footnotesize}
    \\
  }
  \onslide<3->
  \bl{RMS-averaged} %$|Q|^2$
  \begin{footnotesize}
  \begin{align*}
  \rho_{\text{tot},\bigtriangleup} %&= \left[ \int_0^\infty d\ln f\, \f{|{2\tilde{H}_\bigtriangleup(f)\sqrt{f}}|^2}{S_n(f)}\right]^{1/2} 
  &=\f{6}{5}A\, h_0 \rho_c \\%\left[\int_0^{f_\text{ISCO}} d f\, \f{f^{-7/3}}{S_n(f)}\right]^{1/2} \\
  \end{align*}
  \end{footnotesize}
  \\
  %where \begin{small}$\tilde{H}_\bigtriangleup(f)\equiv A\,  \langle |Q|^2_\bigtriangleup\rangle\, h_0 f^{-7/6}= \f{3}{5} A h_0 f^{-7/6}$ \end{small}\\
  \vspace{-5mm}
  Compare with three-L network's \begin{small}$\rho_\text{tot} \approx \tfrac{4}{5}A\, h_0 \left[ \int df \ldots \right]^{1/2} $ \end{small}
  \vspace{10mm}
  %
  %
  \begin{tikzpicture}[overlay,remember picture]
  \node[draw, red,thick,minimum width=3.cm,minimum height=0.9cm] (b) at (-3.2,0.9){};
  \end{tikzpicture}
  %
  % 
\end{frame}

@@

  \begin{columns}
  \begin{column}{0.5\textwidth}  %%<--- here
      \begin{center}
      \onslide<3->\includegraphics[height=4cm]{figs/GW_GRB170817.jpg}
      \end{center}
  \end{column}
  \begin{column}{0.5\textwidth}
  %
  \onslide<6->
  {
    System parameters
    \begin{align*}
     M&=2.73^{+0.04}_{-0.01} M_\odot \\
     M_c &=1.186^{+0.001}_{-0.001}M_\odot  \\ 
     D&=40.7^{+2.36}_{-2.36}\text{Mpc}
    \end{align*}
  }
  \onslide<7->
  {
    \red{Note:} $M_c$ has the best precision: 
   \begin{small}$${\dot{f}} = \f{96}{5}\pi^{8/3} \left(\tfrac{G M_c}{c^3}\right)^{5/3}\, {f^{11/3}}$$ \end{small}
  }
